---
title: "Glendale Water - Census Tracts"
author: "Cameron Tenner"
date: "4/14/2021"
output: html_document
---

```{r setup, include = F}
library(knitr)
opts_chunk$set(echo = T, warning = F, message = F)

library(tidyverse)
library(tigris)
library(censusapi)
library(sf)
library(mapview)
library(plotly)
library(leaflet)
library(tidycensus)

options(
  tigris_class = "sf",
  tigris_use_cache = T 
)
```

Load LA County blockgroups using **tigris** package, import glendale water shapefile using **sf** package
```{r}
blockgroups <-
  block_groups("CA", cb=F, progress_bar=F) %>% 
  filter(
    COUNTYFP == "037" # LA County 
  ) %>% 
  st_as_sf()

glendale_water <- st_read("/Users/ctenner/Desktop/Classes/Github/hr2w_droughts/data/glendale/Water_Districts/Water_Districts.shp") %>% 
  filter(
    AGENCYNAME == "Glendale  City of"
    # grepl("Glendale",
    #       AGENCYNAME),
    # grepl("Water",
    #       AGENCYNAME)
  ) %>% 
  st_as_sf() %>% 
  st_transform(4269)
```

Create subset of Glendale blockgroups, meaning all blockgroups that have any overlap whatsoever with Glendale Water District
```{r}
glendale_blkgps <- 
  blockgroups %>% 
  st_join(glendale_water, join = st_intersects, left = F) %>% 
  st_set_geometry(NULL) %>% 
  left_join(
    blockgroups,
    by = "GEOID"
  ) %>% 
  st_as_sf() %>% 
  select(
    GEOID
  ) %>% 
  filter(
    ! GEOID %in% c("060376512014", "060376214001") # Filters out blockgroups that were selected due to an issue with geometry of glendale water district shapefile
  )
  
```

Find intersection areas using spatial functions and find the % of each blockgroup in the service area
```{r}
intersect_area <- st_area(st_intersection(glendale_blkgps, glendale_water)) # area of intersection for each blockgroup
blkgrp_area <- st_area(glendale_blkgps) # area of each blockgroup

glendale_blkgps <- 
  glendale_blkgps %>% 
  mutate(
    AREA = blkgrp_area,
    INTRSCT_AREA = intersect_area,
    PERC_INTSCT = INTRSCT_AREA / AREA * 100
  )

units(glendale_blkgps$PERC_INTSCT) <- NULL

head(glendale_blkgps)
```

Create interactive map to visualize blockgroup-service area overlap and visually validate % intersections
```{r}
m <- leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = glendale_water %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "red",
    weight = .25,
    fillOpacity = .5,
    label = ~AGENCYNAME
  ) %>% 
  addPolygons(
    data = glendale_blkgps %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "yellow",
    weight = .25,
    fillOpacity = .5,
    label = ~PERC_INTSCT,
    highlightOptions = highlightOptions(
      weight=1.5,
      opacity = 1,
      fillOpacity = .6
    )
  )

m
```

Let's try to remove the spike in blockgroups with close to 0 percent overlap, as they're likely a result of bad geometry in the Glendale Water shapefile 
```{r}
ggplot(glendale_blkgps, aes(x = PERC_INTSCT)) + 
  geom_density() + 
  xlim(0,25) + 
  geom_vline(xintercept = 3, color = "red")

```

3 looks like a decent cutoff, lets see how this looks mapped
```{r}
m <- leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = glendale_water %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "red",
    weight = .25,
    fillOpacity = .5,
    label = ~AGENCYNAME
  ) %>% 
  addPolygons(
    data = glendale_blkgps %>% filter(PERC_INTSCT > 3) %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "yellow",
    weight = .25,
    fillOpacity = .5,
    label = ~PERC_INTSCT,
    highlightOptions = highlightOptions(
      weight=1.5,
      opacity = 1,
      fillOpacity = .6
    )
  )
m
```

Using my knowledge of the area, I can say with confidence that those < 10% intersect should be removed, as the overlap is enitrely unoccupied areas of the blockgroups (ex steep mountainous terrain on side of freeway, open space area, etc)
```{r}
glendale_blkgps_cut <- 
  glendale_blkgps %>% 
  filter(
    PERC_INTSCT >= 10
  )

m <- leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = glendale_water %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "red",
    weight = .25,
    fillOpacity = .5,
    label = ~AGENCYNAME
  ) %>% 
  addPolygons(
    data = glendale_blkgps_cut %>% filter(PERC_INTSCT >= 10) %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "yellow",
    weight = .25,
    fillOpacity = .5,
    label = ~PERC_INTSCT,
    highlightOptions = highlightOptions(
      weight=1.5,
      opacity = 1,
      fillOpacity = .6
    )
  )
m
```

I will use these blockgroups for the remainder of the analysis

Now we load in income data from census table B1901, and join with the glendale blockgroups of interest. We then do some math operations to find the count of households in each income bracket
```{r}
glendale_hh_income <- read_csv("/Users/ctenner/Desktop/Classes/Github/hr2w_droughts/data/glendale/B1901/B1901.csv") %>% 
  separate(
    GEO_ID,
    c("Prefix", "GEOID"),
    "US"
  ) %>% 
  filter(
    GEOID %in% glendale_blkgps_cut$GEOID
  ) %>% 
  select(
    GEOID,
    contains("E"),
    -Prefix,
    -NAME
  )

glendale_data <- glendale_blkgps_cut %>% 
  left_join(glendale_hh_income) %>% 
  st_drop_geometry()

# Convert census data to numeric to allow for calculations 
glendale_data[6:21] <- sapply(glendale_data[6:21], as.numeric)

# Define function for finding num households based on census data and % intersect
fx <- function(x, y) (round(x * y / 100,0))

# Mutate all census columns by multiplying by % intersect 
glendale_data <- glendale_data %>% 
  mutate(
    across(
    6:21,
    ~ fx(.x,y=PERC_INTSCT)
  )) %>% 
  select(
    - B19001_001E
  )

glendale_hh_sum <- tibble(
  hh_income = c("7.5k", "12.5k", "17.5k", "22.5k", "27.5k", "32.5k", "37.5k", "42.5k", "47.5k", "55k", "67.5k", "87.5k", "112.5k", "137.5k", "175k", "250k"),
  count = colSums(glendale_data[,5:20])
  )
```

Define GWD Census Tracts
```{r}
la_tracts <- 
  tracts("CA", cb=F, progress_bar=F) %>% 
  filter(
    COUNTYFP == "037" # LA County 
  ) %>% 
  st_as_sf() 

glen_tracts <- la_tracts[st_intersects(la_tracts, glendale_water, sparse = F),] 

intersect_area <- st_area(st_intersection(glen_tracts, glendale_water)) 
area <- st_area(glen_tracts) 
units(intersect_area) <- NULL
units(area) <- NULL

glen_tracts <- 
  glen_tracts %>% 
  mutate(
    AREA = area,
    INTRSCT_AREA = intersect_area,
    PERC_INTSCT = INTRSCT_AREA / AREA * 100
  ) %>% 
  filter(
    PERC_INTSCT > 10
  )
```

Now load in table of interest from **tidycensus** and subset to tracts identified above
```{r}
income_hh_size <- get_acs(geography="tract",year=2018,state="California",table="B19019") 
glen_income_hh_size <- income_hh_size %>% 
  filter(
    GEOID %in% glen_tracts$GEOID,
    variable != "B19019_001"
  )
```

Now we associate each variable name with the corresponding hh size 
```{r}
sizes <- tibble(
  size=c(1:7),
  varible=unique(glen_income_hh_size$variable)
)

glen_income_hh_size$size <- 0
for(i in 1:nrow(sizes)){
  glen_income_hh_size$size[which(glen_income_hh_size$variable==sizes$varible[i])] <- sizes$size[i]
}
```

Assign income bins to each observation
```{r}
glen_income_hh_size <- 
  glen_income_hh_size %>% 
  mutate(
    income_bracket = case_when(
      estimate < 10000 ~ 7.5,
      estimate >= 10000 & estimate < 15000 ~ 12.5,
      estimate >= 15000 & estimate < 20000 ~ 17.5,
      estimate >= 20000 & estimate < 25000 ~ 22.5,
      estimate >= 25000 & estimate < 30000 ~ 27.5,
      estimate >= 30000 & estimate < 35000 ~ 32.5,
      estimate >= 35000 & estimate < 40000 ~ 37.5,
      estimate >= 40000 & estimate < 45000 ~ 42.5,
      estimate >= 45000 & estimate < 50000 ~ 47.5,
      estimate >= 50000 & estimate < 60000 ~ 55,
      estimate >= 60000 & estimate < 75000 ~ 67.5,
      estimate >= 75000 & estimate < 100000 ~ 87.5,
      estimate >= 100000 & estimate < 125000 ~ 112.5,
      estimate >= 125000 & estimate < 150000 ~ 137.5,
      estimate >= 150000 & estimate < 200000 ~ 175,
      estimate >= 200000 ~ 250
    )
  )
```

Find average hh size for each income bin 
```{r}
income_hh_size_summary <-   
  glen_income_hh_size %>% 
  group_by(income_bracket) %>% 
  summarize(
    hh_size = round(mean(size),2),
    count = n()
  ) 
print.data.frame(income_hh_size_summary)
```

