---
title: "Glendale Water - Census Tracts"
author: "Cameron Tenner"
date: "5/3/2021"
output: html_document
---

```{r setup, include = F}
library(knitr)
opts_chunk$set(echo = T, warning = F, message = F)

library(tidyverse)
library(tigris)
library(censusapi)
library(sf)
library(mapview)
library(plotly)
library(leaflet)
library(tidycensus)

options(
  tigris_class = "sf",
  tigris_use_cache = T 
)
```
Load Santa Cruz County blockgroups using **tigris** package, import Santa Cruz water shapefile using **sf** package
```{r}
bg <-
  block_groups("CA", cb=F, progress_bar=F) %>% 
  filter(
    COUNTYFP == "087" # LA County 
  ) %>% 
  st_as_sf()

sc_water <- st_read("/Users/ctenner/Desktop/Classes/Github/hr2w_droughts/data/santa_cruz/California_Drinking_Water_System_Area_Boundaries/California_Drinking_Water_System_Area_Boundaries.shp") %>% 
  filter(
    WATER_SY_1 == "SANTA CRUZ WATER DEPARTMENT"
  ) %>% 
  st_as_sf() %>% 
  st_transform(4269)
```

Create subset of Glendale blockgroups, meaning all blockgroups that have any overlap whatsoever with Glendale Water District
```{r}
sc_bg <- 
  bg %>% 
  st_join(sc_water, join = st_intersects, left = F) %>% 
  st_set_geometry(NULL) %>% 
  left_join(
    bg,
    by = "GEOID"
  ) %>% 
  st_as_sf() %>% 
  select(
    GEOID
  ) 
  
```

preview the overlap/ selection
```{r}
mapview(list(sc_bg, sc_water))
```
Issues at play: coastal blockgroups (ocean) and rural blockgroups (open space dominated)

cut oceans off of blockgroups for purpose of area calculations
```{r}
ocean <- 
  area_water("CA", 087) %>% 
  filter(
    FULLNAME %in% c("Pacific Ocean", "Monterey Bay")
  ) %>% 
  st_union

sc_bg <- 
  st_difference(sc_bg, ocean)
  
```


Find intersection areas using spatial functions and find the % of each blockgroup in the service area
```{r}
intersect_area <- st_area(st_intersection(sc_bg, sc_water)) # area of intersection for each blockgroup
bg_area <- st_area(sc_bg) # area of each blockgroup

sc_bg <- 
  sc_bg %>% 
  mutate(
    AREA = bg_area,
    INTRSCT_AREA = intersect_area,
    PERC_INT = round(INTRSCT_AREA / AREA * 100, 2)
  )

units(sc_bg$PERC_INT) <- NULL

head(sc_bg)
```

Create interactive map to visualize blockgroup-service area overlap and visually validate % intersections
```{r}
m <- leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = sc_water %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "red",
    weight = .25,
    fillOpacity = .5,
    label = ~WATER_SY_1
  ) %>% 
  addPolygons(
    data = sc_bg %>% st_transform("+proj=longlat +datum=WGS84"),
    fillColor = "yellow",
    weight = .25,
    fillOpacity = .5,
    label = ~PERC_INT,
    highlightOptions = highlightOptions(
      weight=1.5,
      opacity = 1,
      fillOpacity = .6
    )
  )
m
```

Now we load in income data from census table B19001, and join with the Santa Cruz blockgroups of interest. We then do some math operations to find the count of households in each income bracket
```{r}
sc_hh_income <- get_acs(geography="block group",year=2019,state="California",county="Santa Cruz",table="B19001", output="wide") %>% 
  filter(
    GEOID %in% sc_bg$GEOID
  ) %>% 
  select(
    GEOID, 
    contains("E"),
  )

sc_data <- sc_bg %>% 
  left_join(sc_hh_income) %>% 
  st_drop_geometry()

# Define function for finding num households based on census data and % intersect
fx <- function(x, y) (round(x * y / 100,0))

# Mutate all census columns by multiplying by % intersect 
sc_data <- sc_data %>% 
  mutate(
    across(
    6:22,
    ~ fx(.x,y=PERC_INT)
  )) %>% 
  select(
    - B19001_001E
  )

sc_hh_sum <- tibble(
  hh_income = c("7.5k", "12.5k", "17.5k", "22.5k", "27.5k", "32.5k", "37.5k", "42.5k", "47.5k", "55k", "67.5k", "87.5k", "112.5k", "137.5k", "175k", "250k"),
  count = colSums(sc_data[,6:21])
  )
```

Define Santa Cruz Water Census Tracts
```{r}
sc_tracts <- 
  tracts("CA", county = 087, cb=F, progress_bar=F) %>% 
  st_as_sf() 

sc_tracts <- sc_tracts[st_intersects(sc_tracts, sc_water, sparse = F),] 
```

Load in table of interest from **tidycensus** and subset to tracts identified above
```{r}
sc_in_hh_size <- get_acs(geography="tract",year=2018,state="California",table="B19019",county="Santa Cruz") %>% 
  filter(
    GEOID %in% sc_tracts$GEOID,
    variable != "B19019_001"
  )
```

Associate each variable name with the corresponding hh size 
```{r}
sizes <- tibble(
  size=c(1:7),
  varible=unique(sc_in_hh_size$variable)
)

sc_in_hh_size$size <- 0
for(i in 1:nrow(sizes)){
  sc_in_hh_size$size[which(sc_in_hh_size$variable==sizes$varible[i])] <- sizes$size[i]
}
```

Assign income bins to each observation
```{r}
sc_in_hh_size <- 
  sc_in_hh_size %>% 
  mutate(
    income_bracket = case_when(
      estimate < 10000 ~ 7.5,
      estimate >= 10000 & estimate < 15000 ~ 12.5,
      estimate >= 15000 & estimate < 20000 ~ 17.5,
      estimate >= 20000 & estimate < 25000 ~ 22.5,
      estimate >= 25000 & estimate < 30000 ~ 27.5,
      estimate >= 30000 & estimate < 35000 ~ 32.5,
      estimate >= 35000 & estimate < 40000 ~ 37.5,
      estimate >= 40000 & estimate < 45000 ~ 42.5,
      estimate >= 45000 & estimate < 50000 ~ 47.5,
      estimate >= 50000 & estimate < 60000 ~ 55,
      estimate >= 60000 & estimate < 75000 ~ 67.5,
      estimate >= 75000 & estimate < 100000 ~ 87.5,
      estimate >= 100000 & estimate < 125000 ~ 112.5,
      estimate >= 125000 & estimate < 150000 ~ 137.5,
      estimate >= 150000 & estimate < 200000 ~ 175,
      estimate >= 200000 ~ 250
    )
  )
```

Find average hh size for each income bin 
```{r}
income_hh_size_summary <-   
  sc_in_hh_size %>% 
  group_by(income_bracket) %>% 
  summarize(
    hh_size = round(mean(size),2),
    count = n()
  ) 
print.data.frame(income_hh_size_summary)
```
